---
title: "Processing flight data tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(telprep)
library(rlist)
```

This vignette will show you how the telprep package works and how it can be used to process data from telemetry flights. The program is designed to read in raw data, filter out erroneous signals, and help determine the living status of fish.


### Program Workflow:

* Step 1 -- Read the raw data into R 
* Step 2 -- If channels or dates were misprogrammed, make corrections
* Step 3 -- Combine the contents of the data in the folder 
* Step 4 -- Determine the best locations and flag false detections
* Step 5 -- Determine the living status of the fish
* Step 6 -- Add the results to an existing database


### Step 1 -- Read the raw data into R
To use the telprep package, stick the telemetry flight txt files that you want to process into a folder. You can name the files however you choose; however, the names should be recognizable as they will be used later on. Find the path of the folder on your system and use the following lines of code to read the data into R:

```{r eval=FALSE}
folder_path <- "D:/Jordy/telprep/telprep2/data/" # Use your folder path here
raw_data <- read.flight.data(folder_path, remove_999 = T, mort_sig = T)
```

```{r include=FALSE}
load("raw_data")
```
